\HeaderA{Hjorth}{Hjorth Distribution}{Hjorth}
\aliasA{dhjorth}{Hjorth}{dhjorth}
\aliasA{hhjorth}{Hjorth}{hhjorth}
\aliasA{phjorth}{Hjorth}{phjorth}
\aliasA{qhjorth}{Hjorth}{qhjorth}
\aliasA{rhjorth}{Hjorth}{rhjorth}
\keyword{distribution}{Hjorth}
\begin{Description}\relax
These functions provide information about the Hjorth
distribution with location parameter equal to \code{m}, dispersion equal
to \code{s}, and family parameter equal to \code{f}: density,
cumulative distribution, quantiles, log hazard, and random generation.

The Hjorth distribution has density
\deqn{
f(y) = (1+\sigma y)^{-\nu/\sigma} \exp(-(y/\mu)^2/2)
(\frac{y}{\mu^2}+\frac{\nu}{1+\sigma y})}{f(y) = (1+s y)^(-f/s) exp(-(y/m)^2/2) (y/m^2+f/(1+s y))}

where \eqn{\mu}{m} is the location parameter of the distribution,
\eqn{\sigma}{s} is the dispersion, and \eqn{\nu}{f} is the family
parameter.
\end{Description}
\begin{Usage}
\begin{verbatim}
dhjorth(y, m, s, f, log=FALSE)
phjorth(q, m, s, f)
qhjorth(p, m, s, f)
hhjorth(y, m, s, f)
rhjorth(n, m, s, f)
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{y}] vector of responses.
\item[\code{q}] vector of quantiles.
\item[\code{p}] vector of probabilities
\item[\code{n}] number of values to generate
\item[\code{m}] vector of location parameters.
\item[\code{s}] vector of dispersion parameters.
\item[\code{f}] vector of family parameters.
\item[\code{log}] if TRUE, log probabilities are supplied.
\end{ldescription}
\end{Arguments}
\begin{Author}\relax
J.K. Lindsey
\end{Author}
\begin{Examples}
\begin{ExampleCode}
dhjorth(5, 5, 5, 2)
phjorth(5, 5, 5, 2)
qhjorth(0.8, 5, 5, 2)
rhjorth(10, 5, 5, 2)
\end{ExampleCode}
\end{Examples}

