---
title: "Untitled"
author: "Kevin O Brien"
date: "16 December 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Package ‘bssn’
March 17, 2016
Type Package
Title Birnbaum-Saunders Model Based on Skew-Normal Distribution
Version 0.7
Date 2016-03-16
Author Rocio Paola Maehara <rmaeharaa@gmail.com> and Luis Benites
Sanchez <lbenitesanchez@gmail.com>
Maintainer Rocio Paola Maehara <rmaeharaa@gmail.com>
Imports sn
Description It provides the density, distribution function, quantile function, random number generator,
reliability function, failure rate, likelihood function,
moments and EM algorithm for Maximum Likelihood estimators, also empirical quantile
and generated envelope for a given sample, all this for the three parameter
Birnbaum-Saunders model based on Skew-Normal Distribution.
Additionally, it provides the random number generator for the mixture of Birnbaum-
Saunders model based on Skew-Normal distribution.
License GPL (>= 2)
Repository CRAN
NeedsCompilation no
Date/Publication 2016-03-17 23:19:57
R topics documented:
bssn-package . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 2
bssn . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 3
EMbssn . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 6
momentsbssn . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 8
ozone . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 9
reliabilitybssn . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 10
Index 12
1
2 bssn-package
bssn-package Birnbaum-Saunders model based on Skew-Normal distribution
Description
It provides the density, distribution function, quantile function, random number generator, reliability
function, failure rate, likelihood function, moments and EM algorithm for Maximum Likelihood
estimators, also empirical quantile and generated envelope for a given sample, all this for the three
parameter Birnbaum-Saunders model based on Skew-Normal Distribution. Additionally, it provides
the random number generator for the mixture of Birbaum-Saunders model based on Skew-Normal
distribution.
Details
Package: bssn
Type: Package
Version: 0.6
Date: 2015-09-07
License: GPL (>=2)
Author(s)
Rocio Maehara <rmaeharaa@gmail.com> and Luis Benites <lbenitesanchez@gmail.com>

### References

Vilca, Filidor; Santana, L. R.; Leiva, Victor; Balakrishnan, N. (2011). Estimation of extreme percentiles
in Birnbaum Saunders distributions. Computational Statistics & Data Analysis (Print), 55,
1665-1678.
Santana, Lucia; Vilca, Filidor; Leiva, Victor (2011). Influence analysis in skew-Birnbaum Saunders
regression models and applications. Journal of Applied Statistics, 38, 1633-1649.
See Also
bssn, EMbssn, momentsbssn, ozone, reliabilitybssn


bssn Birnbaum-Saunders model based on Skew-Normal distribution

#### Description

It provides the density, distribution function, quantile function, random number generator, likelihood
function, moments and EM algorithm for Maximum Likelihood estimators for a given sample, all this for the three parameter Birnbaum-Saunders model based on Skew-Normal Distribution. Also, we have the random number generator for the mixture of Birbaum-Saunders model based on
Skew-Normal distribution. Finally, the function mmmeth() is used to find the initial values for the
parameters alpha and beta using modified-moment method.

### Usage
<pre><code>

dbssn(ti, alpha=0.5, beta=1, lambda=1.5)
pbssn(q, alpha=0.5, beta=1, lambda=1.5)
qbssn(p, alpha=0.5, beta=1, lambda=1.5)
rbssn(n, alpha=0.5, beta=1, lambda=1.5)
rmixbssn(n,alpha,beta,lambda,pii)
mmmeth(ti)

</code></pre>
### Arguments 

* ``ti``: vector of observations.
* ``q``: vector of quantiles.
* ``p``: vector of probabilities.
* ``n``: number of observations.
* ``alpha``: shape parameter.
* ``beta``: scale parameter.
* ``lambda``: skewness parameter.

pii Are weights adding to 1. Each one of them (alpha, beta and lambda) must be
a vector of length g if you want to generate a random numbers from a mixture
distribution BSSN.
Details
If alpha, sigma or lambda are not specified they assume the default values of 0.5, 1 and 1.5, respectively,
belonging to the Birnbaum-Saunders model based on Skew-Normal distribution denoted
by BSSN(0:5; 1; 1:5).
As discussed in Filidor et. al (2011) we say that a random variable T is distributed as an BSSN with
shape parameter  > 0, scale parameter  > 0 and skewness parameter  in R, if its probability
density function (pdf) is given by
f(t) = 2(a(t; ; ))(a(t; ; ))A(t; ; ); t > 0
where (:) and (:) are the standard normal density and cumulative distribution function respectively.
Also a(t; ; ) = (1=)(
p
t= 􀀀
p
=t) and A(t; ; ) = t􀀀3=2(t + )=(21=2)
4 bssn

### Value

dbssn gives the density, pbssn gives the distribution function, qbssn gives the quantile function,
rbssn generates a random sample and rmixbssn genrates a mixture random sample.
The length of the result is determined by n for rbssn, and is the maximum of the lengths of the
numerical arguments for the other functions dbssn, pbssn and qbssn.
Author(s)
Rocio Maehara <rmaeharaa@gmail.com> and Luis Benites <lbenitesanchez@gmail.com>

### References

Vilca, Filidor; Santana, L. R.; Leiva, Victor; Balakrishnan, N. (2011). Estimation of extreme percentiles
in Birnbaum Saunders distributions. Computational Statistics & Data Analysis (Print), 55,
1665-1678.
Santana, Lucia; Vilca, Filidor; Leiva, Victor (2011). Influence analysis in skew-Birnbaum Saunders
regression models and applications. Journal of Applied Statistics, 38, 1633-1649.
See Also
EMbssn, momentsbssn, ozone, reliabilitybssn

### Examples
```{r}

## Not run:
## Let's plot an Birnbaum-Saunders model based on Skew-Normal distribution!
## Density
sseq <- seq(0,3,0.01)
dens <- dbssn(sseq,alpha=0.2,beta=1,lambda=1.5)
plot(sseq, dens,type="l", lwd=2,col="red", xlab="x", ylab="f(x)", main="BSSN Density function")
# Differing densities on a graph
# positive values of lambda
y <- seq(0,3,0.01)
f1 <- dbssn(y,0.2,1,1)
f2 <- dbssn(y,0.2,1,2)
f3 <- dbssn(y,0.2,1,3)
f4 <- dbssn(y,0.2,1,4)
den <- cbind(f1,f2,f3,f4)
matplot(y,den,type="l", col=c("deepskyblue4", "firebrick1", "darkmagenta", "aquamarine4"), ylab
="Density function",xlab="y",lwd=2,sub="(a)")
legend(1.5,2.8,c("BSSN(0.2,1,1)", "BSSN(0.2,1,2)", "BSSN(0.2,1,3)","BSSN(0.2,1,4)"),
col = c("deepskyblue4", "firebrick1", "darkmagenta", "aquamarine4"), lty=1:4,lwd=2,
seg.len=2,cex=0.8,box.lty=0,bg=NULL)
#negative values of lambda
bssn 5
y <- seq(0,3,0.01)
f1 <- dbssn(y,0.2,1,-1)
f2 <- dbssn(y,0.2,1,-2)
f3 <- dbssn(y,0.2,1,-3)
f4 <- dbssn(y,0.2,1,-4)
den <- cbind(f1,f2,f3,f4)
matplot(y,den,type="l", col=c("deepskyblue4", "firebrick1", "darkmagenta", "aquamarine4"),
ylab ="Density function",xlab="y",lwd=2,sub="(a)")
legend(1.5,2.8,c("BSSN(0.2,1,-1)", "BSSN(0.2,1,-2)","BSSN(0.2,1,-3)", "BSSN(0.2,1,-4)"),
col=c("deepskyblue4","firebrick1", "darkmagenta","aquamarine4"),lty=1:4,lwd=2,seg.len=2,
cex=1,box.lty=0,bg=NULL)
## Distribution Function
sseq <- seq(0.1,6,0.05)
df <- pbssn(q=sseq,alpha=0.75,beta=1,lambda=3)
plot(sseq, df, type = "l", lwd=2, col="blue", xlab="x", ylab="F(x)",
main = "BSSN Distribution function")
abline(h=1,lty=2)
#Inverse Distribution Function
prob <- seq(0,1,length.out = 1000)
idf <- qbssn(p=prob,alpha=0.75,beta=1,lambda=3)
plot(prob, idf, type="l", lwd=2, col="gray30", xlab="x", ylab =
expression(F^{-1}~(x)), mgp=c(2.3,1,.8))
title(main="BSSN Inverse Distribution function")
abline(v=c(0,1),lty=2)
#Random Sample Histogram
sample <- rbssn(n=10000,alpha=0.75,beta=1,lambda=3)
hist(sample,breaks = 70,freq = FALSE,main="")
title(main="Histogram and True density")
sseq <- seq(0,8,0.01)
dens <- dbssn(sseq,alpha=0.75,beta=1,lambda=3)
lines(sseq,dens,col="red",lwd=2)
##Random Sample Histogram for Mixture of BSSN
alpha=c(0.55,0.25);beta=c(1,1.5);lambda=c(3,2);pii=c(0.3,0.7)
sample <- rmixbssn(n=1000,alpha,beta,lambda,pii)
hist(sample$y,breaks = 70,freq = FALSE,main="")
title(main="Histogram and True density")
temp <- seq(min(sample$y), max(sample$y), length.out=1000)
lines(temp, (pii[1]*dbssn(temp, alpha[1], beta[1],lambda[1]))+(pii[2]*dbssn(temp, alpha[2]
, beta[2],lambda[2])), col="red", lty=3, lwd=3) # the theoretical density
lines(temp, pii[1]*dbssn(temp, alpha[1], beta[1],lambda[1]), col="blue", lty=2, lwd=3)
# the first component
lines(temp, pii[2]*dbssn(temp, alpha[2], beta[2],lambda[2]), col="green", lty=2, lwd=3)
# the second component
```

EMbssn EM Algorithm Birnbaum-Saunders model based on Skew-Normal distribution

#### Description
Performs the EM algorithm for Birnbaum-Saunders model based on Skew-Normal distribution.

### Usage
<pre><code>

EMbssn(ti,alpha,beta,delta,loglik=F,accuracy=1e-8,
show.envelope="FALSE",iter.max=500)

</code></pre>
### Arguments 

ti the response vector of dimension n where n is the total of observations.
alpha,beta,delta
an numeric of initial estimates.
loglik showvalue of the log-likelihood (V) or not (F).
accuracy the convergence maximum error.
show.envelope TRUE or FALSE. Indicates if envelope graph should be built for the fitted model.
Default is FALSE.
iter.max The maximum number of iterations of the EM algorithm

### Value

The function returns a list with 11 elements detailed as

* iter number of iterations.
* alpha estimated shape parameter.
* beta estimated scale parameter.
* lambda estimate skewness parameter.
* SE Standard Error estimates.
* table Table containing the inference for the estimated parameters.
* loglik Log-likelihood value.
* AIC Akaike information criterion.
* BIC Bayesian information criterion.
* HQC Hannan-Quinn information criterion.
* time processing time.

Author(s)
Rocio Maehara <rmaeharaa@gmail.com> and Luis Benites <lbenitesanchez@gmail.com>

### References

Vilca, Filidor; Santana, L. R.; Leiva, Victor; Balakrishnan, N. (2011). Estimation of extreme percentiles
in Birnbaum Saunders distributions. Computational Statistics & Data Analysis (Print), 55,
1665-1678.
Santana, Lucia; Vilca, Filidor; Leiva, Victor (2011). Influence analysis in skew-Birnbaum Saunders
regression models and applications. Journal of Applied Statistics, 38, 1633-1649.
See Also
bssn, EMbssn, momentsbssn, ozone, reliabilitybssn

### Examples
```{r}

## Not run:
#Using the ozone data
data(ozone)
attach(ozone)
#################################
#The model
ti <- dailyozonelevel
#Initial values for the parameters
initial <- mmmeth(ti)
alpha0 <- initial$alpha0ini
beta0 <- initial$beta0init
lambda0 <- 0
delta0 <- lambda0/sqrt(1+lambda0^2)
#Estimated parameters of the model (by default)
est_param <- EMbssn(ti,alpha0,beta0,delta0,loglik=T,
accuracy = 1e-8,show.envelope = "TRUE", iter.max=500)
#ML estimates
alpha <- est_param$res$alpha
beta <- est_param$res$beta
lambda <- est_param$res$lambda
#########################################
#A simple output example
---------------------------------------------------------
Birnbaum-Saunders model based on Skew-Normal distribution
---------------------------------------------------------
8 momentsbssn
Observations = 116
-----------
Estimates
-----------
Estimate Std. Error z value Pr(>|z|)
alpha 1.26014 0.23673 5.32311 0.00000
beta 14.65730 4.01984 3.64624 0.00027
lambda 1.06277 0.54305 1.95706 0.05034
------------------------
Model selection criteria
------------------------
Loglik AIC BIC HQC

### Value
 -542.768 4.705 4.741 4.719
-------
Details
-------
Iterations = 415
Processing time = 0.4283214 secs
Convergence = TRUE
## End(Not run)
momentsbssn Moments for the Birnbaum-Saunders model based on Skew-Normal
distribution
Description
Mean, variance, skewness and kurtosis for the Birnbaum-Saunders model based on Skew-Normal
distribution defined in Filidor et. al (2011).

### Usage
<pre><code>

meanbssn(alpha=0.5,beta=1,lambda=1.5)
varbssn(alpha=0.5,beta=1,lambda=1.5)
skewbssn(alpha=0.5,beta=1,lambda=1.5)
kurtbssn(alpha=0.5,beta=1,lambda=1.5)

</code></pre>
  
### Arguments 

* alpha shape parameter .
* beta scale parameter .
* lambda skewness parameter .


### Value

meanbssn gives the mean, varbssn gives the variance, skewbssn gives the skewness, kurtbssn
gives the kurtosis.
Author(s)
Rocio Maehara <rmaeharaa@gmail.com> and Luis Benites <lbenitesanchez@gmail.com>

### References

Vilca, Filidor; Santana, L. R.; Leiva, Victor; Balakrishnan, N. (2011). Estimation of extreme percentiles
in Birnbaum Saunders distributions. Computational Statistics & Data Analysis (Print), 55,
1665-1678.
Santana, Lucia; Vilca, Filidor; Leiva, Victor (2011). Influence analysis in skew-Birnbaum Saunders
regression models and applications. Journal of Applied Statistics, 38, 1633-1649.
See Also
bssn, EMbssn, momentsbssn, ozone, reliabilitybssn

### Examples
```{r}

## Let's compute some moments for a Birnbaum-Saunders model based on Skew normal Distribution.
# The well known mean, variance, skewness and kurtosis
meanbssn(alpha=0.5,beta=1,lambda=1.5)
varbssn(alpha=0.5,beta=1,lambda=1.5)
skewbssn(alpha=0.5,beta=1,lambda=1.5)
kurtbssn(alpha=0.5,beta=1,lambda=1.5)
```

ozone Daily ozone level measurements
Description
These data correspond to daily ozone level measurements (in ppb = ppmx1000) in New York in
May-September, 1973, from the New York State Department of Conservation.

### Usage
<pre><code>

data(ozone)
Format
ozone is a data frame with 116 cases (rows).
Details
For a complete description of various distributions applied to data concentration of air pollutants see Gokhale and Khare (2004).
10 reliabilitybssn

#### Source
Leiva, V., Barros, M., Paula, G. e Sanhueza, A. (2007). Generalized BirnbaumSaunders distribution
applied to air pollutant concentration. Environmetrics, 19, 235-249.
Nadarajah, S. (2007). A truncated inverted beta distribution with application to air pollution data.
Stoch. Environ. Res. Risk. Assess., 22, 285-289.
Gokhale, S. e Khare, M. (2004) A review of deterministic, stochastic and hybrid vehicular exhaust
emission models International. J. Transp. Manag., 2, 59-74.
reliabilitybssn Reliability Function for the Birnbaum-Saunders model based on Skew-
Normal distribution

#### Description
Two useful descriptors in reliability analysis are the reliability function (rf), and the failure rate (fr)
function or hazard function. For a non-negative random variable t with pdf f(t) (and cdf F(t)),
its distribution can be characterized equally in terms of the rf, or of the fr, which are respectively
defined by R(t) = 1 􀀀 F(t), and h(t) = f(t)=R(t), for t > 0,and 0 < R(t) < 1.

### Usage
<pre><code>

Rebssn(ti,alpha=0.5,beta=1,lambda=1.5)
Fbssn(ti,alpha=0.5,beta=1,lambda=1.5)

</code></pre>
### Arguments 

ti dataset.
alpha shape parameter .
beta scale parameter .
lambda skewness parameter .

### Value

Rbssn gives the reliability function, Fbssn gives the failure rate or hazard function.
Author(s)
Rocio Maehara <rmaeharaa@gmail.com> and Luis Benites <lbenitesanchez@gmail.com>

### References

Leiva, V., Vilca-Labra, F. E., Balakrishnan, N. e Sanhueza, A. (2008). A skewed sinh-normal
distribution and its properties and application to air pollution. Comm. Stat. Theoret. Methods.
Submetido.
Guiraud, P., Leiva, V., Fierro, R. (2009). A non-central version of the Birnbaum-Saunders distribution
for reliability analysis. IEEE Transactions on Reliability 58, 152-160.
reliabilitybssn 11
See Also
bssn, EMbssn, momentsbssn, ozone, Rebssn

### Examples
```{r}

## Let's compute some realiability functions for a Birnbaum-Saunders model based on
## Skew normal Distribution for different values of the shape parameter.
ti <- seq(0,2,0.01)
f1 <- Rebssn(ti,0.75,1,1)
f2 <- Rebssn(ti,1,1,1)
f3 <- Rebssn(ti,1.5,1,1)
f4 <- Rebssn(ti,2,1,1)
den <- cbind(f1,f2,f3,f4)
matplot(ti,den,type="l", col=c("deepskyblue4","firebrick1","darkmagenta","aquamarine4"),
ylab="S(t)", xlab="t",lwd=2)
legend(1.5,1,c(expression(alpha==0.75), expression(alpha==1), expression(alpha==1.5),
expression(alpha==2)),col= c("deepskyblue4","firebrick1","darkmagenta","aquamarine4"),
lty=1:4,lwd=2,seg.len=2,cex=0.9,box.lty=0,bg=NULL)
## Let's compute some hazard functions for a Birnbaum Saunders model based on
## Skew normal Distribution for different values of the skewness parameter.
ti <- seq(0,2,0.01)
f1 <- Fbssn(ti,0.5,1,-1)
f2 <- Fbssn(ti,0.5,1,-2)
f3 <- Fbssn(ti,0.5,1,-3)
f4 <- Fbssn(ti,0.5,1,-4)
den <- cbind(f1,f2,f3,f4)
matplot(ti,den,type = "l", col = c("deepskyblue4","firebrick1", "darkmagenta", "aquamarine4"),
ylab = "h(t)" , xlab="t",lwd=2)
legend(0.1,23, c(expression(lambda==-1), expression(lambda==-2), expression(lambda == -3),
expression(lambda == -4)), col = c("deepskyblue4", "firebrick1", "darkmagenta","aquamarine4"),
lty=1:4,lwd=2,seg.len=2,cex=0.9,box.lty=1,bg=NULL)
```
